### 2019-01-10 To_Do.Rmd
USA preferences done

---
title: "To Do"
author: ""
date: "29/10/2018"
output: html_document
---

# Data

## Trends

[ ] Bartomeus PNAS estimates (Nacho) 
[X] Scheper PNAS estimates (MA)
[ ] UE red list (via @ropensci; MA)
[X] SPAIN trends (MA: species with Brain size) 
[ ] Trends: Nacho or Cristina

## traits

[ ] LHT Bartomeus PNAS + Scheper PNAS + Traitbase (Var) Nacho
[X] Brain: MA (Spain + ND + US Ferran + US Parker): MA
[X] Habitats USA: MA 
    [ ] Others ¿?: MA
[ ] Exotic use ¿? (Nacho)
[X] Innovation (some in MA bbdd)
[ ] Invasive species in US. (Nacho)

## Phylogeny

[ ] Hedtke et al 2013? + politomies
[ ] Build one as Ferran did?

# Models

- Regresion Trees (see Sol et al 2012 Science)
- Quantile regresions

# Refs:

- Mamifers. Manuela gonzalez??? big-worst
- Birds. Invasion proceses (Sol et al 2012),... 
- Reptiles. Invasion also related with Brain size.