### 2022-10-25 Scripts/Figure_4_Map_Points/Map.Rmd
edits

---
title: ""
output:
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12}

#Plot maps with density of bee occurrences

#Europe----
#Load libraries
library(giscoR) #map of europe
library(ggplot2)
library(dplyr)
library(readr)
library(sf)
library(patchwork)
library(rnaturalearth)

#Load europe map
Europe <- gisco_get_countries(region = "Europe")
#Load countries filtered
euro_map <- st_read("../../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
#Load data with records filtered
all_long_lat <- read.table("../../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")

#Plot Europe map with counties of interest highlighted
p1 = ggplot(Europe) +
geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
panel.background = element_rect(fill = "white"))

p2 = ggplot(Europe) +
geom_sf(fill="gray70",color = "black", alpha = 1,size=0.2)+
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8)) +
coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
theme(panel.background = element_rect(fill = "aliceblue"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("A") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(), 
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 12),
plot.title = element_text(face="bold"),
legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(40, 60, by = 6)) +
scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))

#Insert general plot within specific plot
panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)

#USA-----
#Load libraries

#Load data
usa_states_1 <- read.csv("../../Data/Usa_data/usa_states.csv")
ny_metropolitan_1 <- read.csv("../../Data/Usa_data/ny_metropolitan.csv")
all_long_lat <- read.table("../../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")

#Load maps
canada <- ne_states(country = "canada", returnclass = "sf") %>%
st_as_sf(crs = 4326)

usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
st_as_sf(crs = 4326)

p1 = ggplot(canada) + 
geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
panel.background = element_rect(fill = "white")) +
geom_map(data = ny_metropolitan_1, 
map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
size = 0.1) +
geom_map(data = usa_states_1, map = usa_states_1,
aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
size = 0.1) +
coord_sf(xlim = c(-86, -62), ylim = c(25, 48))

p2 = ggplot(canada) + 
geom_sf(fill="gray70",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray70",color = "black", alpha = 1,size=0.2) +
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 3, 6)) +
coord_sf(xlim = c(-82, -68), ylim = c(33, 47))  +
theme(panel.background = element_rect(fill = "aliceblue"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("B") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(),
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 12),
plot.title = element_text(face="bold"),
legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(33, 47, by = 4)) +
scale_x_continuous(breaks = seq(-82, -68, by = 4)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))

#Insert general plot within specific plot
panel2 = p2 + inset_element(p1, left = 0, bottom = -0.885, right = 0.35, top = 1.2, ignore_tag = TRUE)

library(patchwork)
panel1 + panel2

```