### 2022-12-06 Manuscript/trial.Rmd
figure 3 update

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
date: "2022-12-06"
---
```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=4, fig.width=4, fig.cap="\\textbf{Figure 3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban). The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}

#Histogram of habitat preferences
#Load libraries
library(tidyverse)

#Load preferences
preferences_europe = read_csv("../Data/Europe_data/preferences_europe.csv") 
preferences_usa = read_csv("../Data/Usa_data/preferences_usa.csv")

#Merge datasets
preferences = bind_rows(preferences_europe, preferences_usa)

#Convert to long to model everything at the same time
long_data = preferences %>% gather(Habitat, Preference, 2:4, -c(Species))

#Relevel for plotting
long_data = long_data %>% 
mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))

labels.minor <- c("Avoidance", rep("", times = 2), "Preference","")

long_data$Habitat <- factor(long_data$Habitat, levels = c("Natural", "Agricultural", "Urban"),labels = c("A) Natural", "B) Agricultural", "C) Urban"))

long_data = long_data %>% 
mutate(Preference_colour = case_when(Preference <= 0.2 ~ "A", Preference >= 0.8 ~ "B", T ~ "C"))

long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Avoidance", "No preference", "Preference"))

ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
    xlab("Habitat preference") +
    ylab("Frequency") +
    facet_wrap(~ Habitat) +
    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
    theme_bw() +   
    theme(axis.title = element_text(face = "bold"),
    panel.border=element_rect(size=1),
    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
    panel.grid.minor=element_blank())+
    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
                       labels = c(0,0.2,0.4,0.6,0.8,1))

```